Learning constraints in tabular data

### 2016-05-19 Paper/ECAI/tabular.tex
Fig: add all constraints
Rewrite algorithm intro

  \begin{subfigure}{.70\textwidth}
  \begin{center}
    \includegraphics[width=1\textwidth]{figures/Demo.png}
  \end{center}
  \vspace{-10pt}
  \caption{Example spreadsheet (black words and numbers only). Green borders indicate headerless tables; the blue elements are added to clarify some constraints and relations. \tias{TODO: headers not in green boxes} \sergey{show blocks? orientation only over blocks, not over tables?}\tias{Use specific constraint names, e.g. 'SUM-row' above Total in T1}}
  \label{fig:main_example}
\end{subfigure}
\hfill
\begin{subfigure}{.30\textwidth}
  {\small
    \begin{align*}
      &~\ecseries{\range{T_{1}}{\rangeall}{1}} \\
%
      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}} \\
%
      &~\ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecsumc{\range{T_{6}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
%
      &~\ecsumr{\range{T_{1}}{\rangeall}{7}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
%
      &~\ecaggc{\textit{AVERAGE}}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggc{\textit{MAX}}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggc{\textit{MIN}}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggif{\textit{SUM}}{\range{T_{1}}{\rangeall}{10}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
%
      &~\ecaggif{\textit{MAX}}{\range{T_{1}}{\rangeall}{11}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
%
      &~\eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}} \\
%
      &~\range{T_{6}}{\rangeall}{4} = PREV(\range{T_{6}}{\rangeall}{4}) + \range{T_{6}}{\rangeall}{2} - \range{T_{6}}{\rangeall}{3}
    \end{align*}}
  \vspace{-10pt}
  \caption{Constraints present in the running example (left) \tias{Add ALL, including non-functional and duplicates}.}
  \label{fig:sol_example}
\end{subfigure}
\end{figure*}

@@

\paragraph{Automatic detection}
As mentioned above, automatic detection of headers and tables is a complex task.
Therefore, we define a restricted setting in which the task becomes easy enough to be automated.
We assume that 1) tables are rectangular blocks of data not containing any empty cells; and 2) tables are separated by at least one empty cell from other tables.
Additionally, if the first row contains only textual values it treated as a header, otherwise, if the first column consists of only textual data is will be the header.
If a header row (column) is detected, it is removed from the table and the orientation of the table is fixed to columns (rows), otherwise, we assume there is no header and the orientation is not fixed.

@@

\paragraph{Visual}
Since the specification of tables is usually easy for humans, we propose a second approach which aims to decrease the complexity of specifying tables by allowing users to indicate tables using a visual tool (Figure~\ref{fig:visual_tool}).
Users select tables excluding the header and optionally specify an orientation.
The tool then generates the table definitions automatically.
\\\\
Aside from extending the automatic detection to be more applicable, one could combine the visual and automatic detection to let the user approve or correct detected table.

@@

\newcommand{\temps}{\ensuremath{S}}
\subsection{Algorithm} \label{sec:algo}
Our method solves the \tcl problem by checking for each template what combination of groups can satisfy the signature, and which specific subgroups satisfy the definition.

The main challenge is to avoid having to check each possible combination of group and subgroup on each constraint template, as well as developing an \textit{extensible} method in which new constraint templates can be plugged in with relative ease. %We will use constraint satisfaction technology to partly answer this challenge.
To answer this challenge we separate checking the signature of a constraint template from checking the definition of a constraint template. Furthermore, we use constraint satisfaction technology to jointly search for and test the signature over all group combinations.

The pseudo-code of the main method is shown in Algorithm~\ref{algo:tcl}; it consists of three key components: 1) templates can depend on each other, so they are ordered to avoid having to verify the same template multiple times; 2) we generate all combinations of groups that can satisfy the signature using constraint satisfaction techniques; 3) from this, all satisfying subgroups are generated.

@@

%First, constraint templates are ordered such that they can be treated sequentially and each only once.
%Then, for each constraint template~$s \in \temps$ we generate a set~$A$ of \textit{group-assignments} that satisfy the \CSignature of~$s$.
%Every group assignment assigns a group~$G \in \groups$ to every argument of the template.
%In the third step, for every assignments $(G_1, ..., G_n) \in A$ we find all constraints $s(G_1', ... G_n')$ that hold on the data (satisfy the \CFunction), such that every subgroup $G_i' \subseteq G_i$.

%Our Algorithm \ref{algo:tcl} is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}, however, we apply the approach on two levels, generating and testing groups according to their properties and generating and testing subgroups on the actual data.
%We will explain and illustrate the three key steps in detail.

We now describe each of the three components in more detail.

\subsubsection{Template ordering}
As discussed before, a constraint may imply another constraint. This is apparent when looking at the template signatures in Table~\ref{table:constraints}, where one can see that some signatures depend on other constraints, such as $LOOKUP$ depending on $FOREIGNKEY$.

>>>>

  \begin{subfigure}{.70\textwidth}
  \begin{center}
    \includegraphics[width=1\textwidth]{figures/Demo.png}
  \end{center}
  \vspace{-10pt}
  \caption{Example spreadsheet (black words and numbers only). Green borders indicate headerless tables; the blue elements are added to clarify some constraints and relations. \tias{TODO: headers not in green boxes} \sergey{show blocks? orientation only over blocks, not over tables?}\tias{Use specific constraint names, e.g. 'SUM-row' above Total in T1}}
  \label{fig:main_example}
\end{subfigure}
\hfill
\begin{subfigure}{.30\textwidth}
  {\footnotesize
    \begin{align*}
      % Table 1
      &~\ecseries{\range{T_{1}}{\rangeall}{1}} \\
%
      &~\ecperm{\range{T_{1}}{\rangeall}{1}}, \ecperm{\range{T_{1}}{\rangeall}{8}} \\
%
      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{3}}, \ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{4}} \\
      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}}, \ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{5}} \\
      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{6}}, \ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{10}} \\
%
      &~\ecsumr{\range{T_{1}}{\rangeall}{7}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
%
      &~\ecaggif{\textit{SUM}}{\range{T_{1}}{\rangeall}{10}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
%
      &~\ecaggif{\textit{MAX}}{\range{T_{1}}{\rangeall}{11}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
%
	  % Table 2
      &~\ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggc{\textit{AVERAGE}}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
      &~\ecaggc{\textit{MAX}}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}},  \\
%
      &~\ecaggc{\textit{MIN}}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
%
	  % Table 4
      &~\eclookup{\range{T_{4}}{\rangeall}{2}}{\range{T_{4}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}} \\
%
      &~\eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}} \\
%
	  % Table 6
      &~\ecsumc{\range{T_{6}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
%
      &~\range{T_{6}}{\rangeall}{4} = PREV(\range{T_{6}}{\rangeall}{4}) + \range{T_{6}}{\rangeall}{2} - \range{T_{6}}{\rangeall}{3}
    \end{align*}}
  \vspace{-10pt}
  \caption{Constraints present in the running example (left), except \textit{ALLDIFFERENT} (16) and \textit{FOREIGNKEY} (5).}
  \label{fig:sol_example}
\end{subfigure}
\end{figure*}

@@

\subsubsection{Automatic detection}
As mentioned above, automatic detection of headers and tables is a complex task.
Therefore, we define a restricted setting in which the task becomes easy enough to be automated.
We assume that 1) tables are rectangular blocks of data not containing any empty cells; and 2) tables are separated by at least one empty cell from other tables.
Additionally, if the first row contains only textual values it treated as a header, otherwise, if the first column consists of only textual data is will be the header.
If a header row (column) is detected, it is removed from the table and the orientation of the table is fixed to columns (rows), otherwise, we assume there is no header and the orientation is not fixed.

@@

\subsubsection{Visual}
Since the specification of tables is usually easy for humans, we propose a second approach which aims to decrease the complexity of specifying tables by allowing users to indicate tables using a visual tool (Figure~\ref{fig:visual_tool}).
Users select tables excluding the header and optionally specify an orientation.
The tool then generates the table definitions automatically.
\\\\
Aside from extending the automatic detection to be more applicable, one could combine the visual and automatic detection to let the user approve or correct detected table.

@@

\newcommand{\temps}{\ensuremath{S}}

\subsection{Algorithm}
\label{sec:algo}
Our method assumes that tables and blocks are given and solves the \tcl~problem by checking for each template: what combination of given (super)blocks may satisfy the signature and which specific (sub)blocks satisfy both the signature and the definition.

The main challenge we aim to address with our approach is to avoid having to check each possible combination of blocks for each constraint template as well as developing an \textit{extensible} method in which new constraint templates can be plugged in with relative ease. %We will use constraint satisfaction technology to partly answer this challenge.
To answer this challenge we separate checking the signature (concerning properties of a block) of a constraint template from checking the definition (concerning the values in a block) of a constraint template.
Furthermore, we use constraint satisfaction technology to jointly search for and test the signature over all combinations of superblocks.

@@

The pseudo-code of the main method is shown in Algorithm~\ref{algo:tcl}; it consists of three key components: 1) templates can depend on each other, so they are ordered to avoid having to verify the same template multiple times; 2) we generate all combinations of superblocks that can satisfy the signature using constraint satisfaction techniques; 3) from this, all satisfying subblocks are generated.
We now describe each of the three components in more detail.

\samuel{MARKER-C}
\subsubsection{Template ordering}
As discussed before, a constraint may imply another constraint. This is apparent when looking at the template signatures in Table~\ref{table:constraints}, where one can see that some signatures depend on other constraints, such as $LOOKUP$ depending on $FOREIGNKEY$.
